---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# factorselect

<!-- badges: start -->
<!-- badges: end -->

An R package for eigenvalue-based estimation of the number of factors 
in approximate factor models. Designed to work when either N or T is 
large, without requiring both dimensions to grow simultaneously.

## Installation

```r
# Development version from GitHub
devtools::install_github("penny4nonsense/factorselect")
```

## Basic Usage

```{r example}
library(factorselect)

# Simulate a factor model with 3 true factors
set.seed(42)
X <- simulate_factor_model(N = 100, T = 200, k = 3)

# Estimate number of factors
select_factors(X, method = "ahn_horenstein")
```

## Methods Implemented

| Method | Reference |
|--------|-----------|
| Eigenvalue Ratio (ER, GR) | Ahn & Horenstein (2013) |
| Information Criteria | Bai & Ng (2002) |
| Edge Distribution Test | Onatski (2009, 2010) |
| ABC Penalty | Alessi, Barigozzi & Capasso (2010) |
| Ratio-Based | Lam & Yao (2012) |

## References

Ahn, S.C. and Horenstein, A.R. (2013). Eigenvalue Ratio Test for the
Number of Factors. *Econometrica*, 81(3), 1203-1227.
