---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# factorselect

An R package for eigenvalue-based estimation of the number of factors
in approximate factor models. Designed to work when either N or T is
large, without requiring both dimensions to grow simultaneously.

## Installation

```r
# Development version from GitHub
devtools::install_github("penny4nonsense/factorselect")
```

## Basic Usage

```{r example}
library(factorselect)

# Simulate a factor model with 3 true factors
X <- simulate_factor_model(N = 100, TT = 200, k = 3, sd = 0.5, seed = 42)

# Estimate number of factors using Ahn & Horenstein (2013)
result <- select_factors(X, method = "ahn_horenstein", kmax = 8)
print(result)
```

## All Methods

```{r all_methods}
# Run all estimators at once
result <- select_factors(X,
                         method = c("ahn_horenstein", "bai_ng", "abc",
                                    "lam_yao", "onatski_2009", "onatski_2010"),
                         kmax   = 8)
print(result)
```

## Scree Plot

```{r plot, fig.height = 4, fig.width = 6}
# Visualize the eigenvalue spectrum
result <- select_factors(X, method = "ahn_horenstein", kmax = 8)
plot(result)
```

## Methods Implemented

| Method | Reference |
|--------|-----------|
| Eigenvalue Ratio (ER, GR) | Ahn & Horenstein (2013) |
| Information Criteria (PC, IC) | Bai & Ng (2002) |
| Tuned Information Criteria | Alessi, Barigozzi & Capasso (2010) |
| Auto-covariance Ratio | Lam & Yao (2012) |
| Edge Distribution Test | Onatski (2009) |
| Edge Distribution Estimator | Onatski (2010) |

## References

Ahn, S.C. and Horenstein, A.R. (2013). Eigenvalue Ratio Test for the
Number of Factors. *Econometrica*, 81(3), 1203-1227.

Bai, J. and Ng, S. (2002). Determining the Number of Factors in
Approximate Factor Models. *Econometrica*, 70(1), 191-221.

Alessi, L., Barigozzi, M. and Capasso, M. (2010). Improved Penalization
for Determining the Number of Factors in Approximate Factor Models.
*Statistics and Probability Letters*, 80, 1806-1813.

Lam, C. and Yao, Q. (2012). Factor Modelling for High-Dimensional Time
Series: Inference for the Number of Factors. *The Annals of Statistics*,
40(2), 694-726.

Onatski, A. (2009). Testing Hypotheses About the Number of Factors in
Large Factor Models. *Econometrica*, 77(5), 1447-1479.

Onatski, A. (2010). Determining the Number of Factors From Empirical
Distribution of Eigenvalues. *The Review of Economics and Statistics*,
92(4), 1004-1016.
